---
title: "01 - The Data"
author: "Lisa Breckels"
date: "15/04/2021"
output: html_document
---

All mass spectrometry proteomics data from this study have been deposited to the [ProteomeXchange](http://proteomecentral.proteomexchange.org/cgi/GetDataset) Consortium via the PRIDE partner repository [1], with the dataset identifier PXD023509. They are also freely and openly available in the [Bioconductor pRolocdata package](https://bioconductor.org/packages/release/data/experiment/html/pRolocdata.html) (≥v1.27.3) [2].

## Accessing the data from the Bioconductor pRolocdata package in R
All analysis in the manuscipt is done in the R programming language, unless otherwise stated using the suite of mass spectrometry spatial proteomics packages `MSnbase`, `pRoloc`, `pRolocGUI` and `pRolocdata`.

The `pRolocdata` package is a R [Bioconductor](http://bioconductor.org/) [experiment package](http://bioconductor.org/packages/release/BiocViews.html#___ExperimentData)([release](http://bioconductor.org/packages/release/data/experiment/html/pRolocdata.html) and [devel](http://bioconductor.org/packages/devel/data/experiment/html/pRolocdata.html) pages) that collects published (mainly) mass spectrometry-based spatial/organelle and protein-complex dataset. Each data is wrapped into a convenient container called a `MSnSet` instance (see the [`MSnbase`](http://www.bioconductor.org/packages/release/bioc/html/MSnbase.html) for details) so work seamlessly with the [`pRoloc`](http://bioconductor.org/packages/release/data/experiment/html/pRolocdata.html) and [`pRolocGUI`](http://bioconductor.org/packages/devel/bioc/html/pRolocGUI.html) software for spatial proteomics data analysis and visualisation. A  

The first step is to install the packages from Bioconductor
```{r setup, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("pRoloc")
BiocManager::install("pRolocdata")
```

Now we can load the packages by calling `library` the loading the data
```{r loaddata}
library("pRoloc")
library("pRolocdata")
data()
```

[1] Perez-Riverol, Y. et al. The PRIDE database and related tools and resources in 2019: improving support for quantification data. Nucleic Acids Res 47, D442-D450, [doi:10.1093/nar/gky1106](doi:10.1093/nar/gky1106) (2019).
[2] Gatto, L., Breckels, L. M., Wieczorek, S., Burger, T. & Lilley, K. S. Mass-spectrometry-based spatial proteomics data analysis using pRoloc and pRolocdata. Bioinformatics 30, 1322-1324, [doi:10.1093/bioinformatics/btu013](doi:10.1093/bioinformatics/btu013) (2014)
